### 2021-01-20 EDA Project 1 NN 6372.Rmd
Add files via upload

---
title: "EDA 6372 Project 1 NN"
author: "Nicole Norelli"
date: "1/20/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Read in data, fix factors, examine NAs
cars = read.csv('/Users/nicolenorelli/Documents/DS6372/Project1Details_2021/data1.csv')
cars$Make = as.factor(cars$Make)
cars$Transmission.Type = as.factor(cars$Transmission.Type)
cars$Driven_Wheels = as.factor(cars$Driven_Wheels)
cars$Vehicle.Size = as.factor(cars$Vehicle.Size)
cars$Vehicle.Style = as.factor(cars$Vehicle.Style)
cars$Engine.Fuel.Type = as.factor(cars$Engine.Fuel.Type)
cars$Model = as.factor(cars$Model)

# Need a closer look at Market.Category
# Might want to split up the multiple categories
# There are a bunch of NAs in there too as "N/A"

sapply(cars,function(x) sum(is.na(x))) 
# 6 in Number.of.Doors, 30 in Engine.Cylinders, 69 in Engine.HP
# most of the cylinder NAs are the Mazda rx-8, which doesn't really have cylinders

summary(cars)
```

```{r}
library(tidyverse)
# A look at MSRP
cars %>% ggplot(aes(x=MSRP)) + geom_histogram()
cars %>% ggplot(aes(x=log(MSRP))) + geom_histogram()
cars %>% filter(MSRP<1000000) %>% ggplot(aes(x=MSRP)) + geom_histogram()
cars %>% filter(MSRP<1000000) %>% ggplot(aes(x=log(MSRP))) + geom_histogram()

# Summary MSRP stats for most Factors
t(aggregate(MSRP~Engine.Cylinders, data=cars, summary))
t(aggregate(MSRP~Engine.Fuel.Type, data=cars, summary))
t(aggregate(MSRP~Vehicle.Size, data=cars, summary))
t(aggregate(MSRP~Transmission.Type, data=cars, summary))
t(aggregate(MSRP~Vehicle.Style, data=cars, summary))
t(aggregate(MSRP~Driven_Wheels, data=cars, summary))
```

```{r}
# Visualizations of MSRP and possible predictor variables
plot(as.factor(cars$Engine.Cylinders), cars$MSRP, xlab="Cylinders",ylab="MSRP")
plot(as.factor(cars$Engine.Cylinders), log(cars$MSRP), xlab="Cylinders",ylab="logMSRP")
p1 = cars %>% ggplot(aes(x=Engine.Fuel.Type,y=MSRP)) + geom_boxplot()
p1 + theme(axis.text.x = element_text(angle=45, hjust=1))
cars %>% ggplot(aes(x=Transmission.Type,y=MSRP)) + geom_boxplot()
cars %>% ggplot(aes(x=Driven_Wheels,y=MSRP)) + geom_boxplot()
cars %>% ggplot(aes(x=Vehicle.Size,y=MSRP)) + geom_boxplot()
p = cars %>% ggplot(aes(x=Vehicle.Style,y=MSRP)) + geom_boxplot()
p + theme(axis.text.x = element_text(angle=45, hjust=1))
```

```{r}
# MSRP correlations with numeric predictors
library(GGally)
cars %>% select(MSRP, Year, Popularity) %>% ggpairs()
cars %>% select(MSRP, Engine.HP, Engine.Cylinders, highway.MPG, city.mpg) %>% ggpairs()
cars %>% select(MSRP, Year, Popularity) %>% ggpairs()
cars %>% filter (!is.na(Engine.HP)) %>% select(MSRP, Engine.HP, Engine.Cylinders, highway.MPG, city.mpg) %>% ggpairs()
# temporarily removed NAs for convenience in this graph

# Multicollinearity examination
pairs(cars[,c(3,5,6,9,13,14,15,16)])
```

```{r}

```